Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-16 manuscript/letter.tex
Changes to diff between disturbance and anomaly
When defining the free-air anomaly, the reduction is applied to g, not
gamma. It's reducing g from P to Q', not raising gamma.

I really don't like the term "geoidal undulation" (it has nothing to do
with waves) or, even worse, "geoid anomaly". N is a distance so "geoidal
height" is better and is what is used in the Physical Geodesy book.

Different gravity anomalies can be calculated,
depending on the corrections applied to them.
These corrections are usually called \textit{gravity reductions}.
The Free-air anomaly, for example, may be defined as
follows \citep{blakely1996, hofmann-wellenhof-moritz2005}:
\begin{equation}
\Delta g_{P}^{F}
= g_{P} - \left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} \, H_{P} \right) \: ,
\label{eq:free-air-anomaly}
\end{equation}
where $\frac{\partial \gamma}{\partial h} \approx -0.3086$ mGal/m is the
derivative of the normal gravity with respect to the geometric height $h$
and $H_{P}$ is the orthometric height $H$ (Fig. \ref{fig:surfaces}) at the
point $P$.
Notice that the term between parenthesis represents an
approximation of the normal gravity at the point $Q^{\prime}$ on the
Geoid. By using a similar approach, we can approximate the normal gravity
$\gamma_{P}$, at a point $P$ on the Earth's surface, so that the
gravity disturbance $\delta g_{P}$ (eq. \ref{eq:gravity-disturbance})
can be rewritten as follows:
\begin{equation}
\delta g_{P} \approx g_{P} -
\left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} h \right) \: .
\label{eq:gravity-disturbance-approx}
\end{equation}
By inspection, one easily concludes that the absolute difference
between the approximated gravity disturbance $\delta g_{P}$
(eq. \ref{eq:gravity-disturbance-approx}), at a point $P$ on or close to
the Earth's surface, and the Free-air anomaly
$\Delta g_{P}^{F}$ (eq. \ref{eq:free-air-anomaly}), is given by:
\begin{equation}
\vert \delta g_{P} - \Delta g^{F}_{P} \vert \approx
\vert \frac{\partial \gamma}{\partial h} N \vert \: ,
\label{eq:difference-disturbance-free-air}
\end{equation}
where $N \approx h - H$ is the geoidal undulation (Fig. \ref{fig:surfaces}).
This approximation assumes that the Geoid and the surface of
the reference ellipsoid are approximately parallel at $P$
and the surrounding area.
We know empirically that the geoidal undulation $N$ in the world
is $\approx \pm 1$ m on the oceans and reaches a
maximum absolute value of $\approx 120$ m \citep[e.g.,][]{torge2012,
sanso_sideris2013}. In Brazil, for example, the geoidal undulation
reaches $\approx \pm 30$ m \citep{ibge_mapgeo2015}. Consequently, the
maximum absolute differences between gravity disturbance and
Free-air anomaly, according to eq. \ref{eq:difference-disturbance-free-air},
is $\approx 9.258$ mGal.
The approximations defined by eqs \ref{eq:gravity-disturbance-approx}
and \ref{eq:difference-disturbance-free-air} are commonly used
in geodesy to define the \textit{fundamental equation of physical
geodesy} \citep{hofmann-wellenhof-moritz2005}.

>>>>

Different approximations for the gravity anomaly can be calculated
depending on the corrections applied to them.
These corrections are usually called \textit{gravity reductions}
and their aim is to approximate the magnitude of gravity at the geoid
$g_{Q^\prime}$ from the magnitude of gravity $g_P$ measured at a point $P$
on or above the surface of the Earth.
The \textit{free-air anomaly}, for example, can be defined as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

\begin{equation}
\Delta g^{F}
= \left(g_{P} - \frac{\partial \gamma}{\partial h} \, H_{P} \right)
- \gamma_{Q} \: ,
\label{eq:free-air-anomaly}
\end{equation}

\noindent
where $\partial \gamma/\partial h \approx -0.3086$ mGal/m is the
derivative of the normal gravity with respect to the geometric height $h$
and $H_{P}$ is the orthometric height (i.e., with respect to the geoid)
of point $P$ (Fig. \ref{fig:surfaces}).
The term between parenthesis represents an
approximation of the gravity at the point $Q^{\prime}$ on the geoid.
Similarly, we can approximate the normal gravity $\gamma_{P}$
at a point $P$ on the Earth's surface
from the normal gravity at the ellipsoid $\gamma_{Q}$,
so that the gravity disturbance $\delta g_{P}$
(eq. \ref{eq:gravity-disturbance}) can be rewritten as:

\begin{equation}
\delta g_{P} \approx g_{P} -
\left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} h_P \right) \: .
\label{eq:gravity-disturbance-approx}
\end{equation}

\noindent
where $h_P$ is the geometric height of point P.
Thus, the absolute difference between
the approximated gravity disturbance (eq. \ref{eq:gravity-disturbance-approx}),
at a point $P$ on or close to the Earth's surface,
and the Free-air anomaly (eq. \ref{eq:free-air-anomaly}) is:

\begin{equation}
\left\vert \delta g_{P} - \Delta g^{F} \right\vert \approx
\left\vert \frac{\partial \gamma}{\partial h} N \right\vert \: ,
\label{eq:difference-disturbance-free-air}
\end{equation}

\noindent
where $N \approx h - H$ is the geoidal height (Fig. \ref{fig:surfaces}).
This approximation assumes that the geoid and the surface of
the reference ellipsoid are close to being parallel at $P$
and the surrounding area.
Eqs \ref{eq:gravity-disturbance-approx}
and \ref{eq:difference-disturbance-free-air} are commonly used
in geodesy to define the \textit{fundamental equation of physical
geodesy} \citep{hofmann-wellenhof-moritz2005}.

We know empirically that $N$ is $\approx \pm 1$ m on the oceans
and reaches a maximum absolute value of $\approx 120$ m
\citep[e.g.,][]{torge2012, sanso_sideris2013}.
In Brazil, for example, the geoidal height
varies between $\approx \pm 30$ m \citep{ibge_mapgeo2015}.
Consequently,
the maximum absolute difference between the gravity disturbance and
the free-air anomaly in Brazil,
according to eq. \ref{eq:difference-disturbance-free-air},
is $\approx 9.258$ mGal.
